---
title: "Assignment1_Hintz"
author: "Hintz A."
date: "2026-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
```

```{css, echo = FALSE}
body {
  text-align: justify;
  font-size: 16px;
}
```


# Exercise 1

1. Using your own words, describe a randomized complete block design (RCBD)
in 3-5 sentences. Then write out a statistical model (using mathematical notation)
that could be fitted to data generated by an RCBD.


The randomized complete block design (RCBD) is a structure that is used when we do not have homogeneous experimental units, which is a requirement to fit an appropriate model. Therefore, we divide the experimental units into homogeneous groups (blocks), where each treatment appears at least once in each block. In this way, we can control the local variability. When fitting an RCBD, we assume that the errors are independent and follow a normal distribution with mean 0 and constant variance. When we do not use blocks, the error increases, and consequently the confidence intervals become wider, making it more difficult to detect differences among treatments, that is, failing to reject the null hypothesis $H_0$ when at least one treatment has a significant effect.

The RCBD model can be written as

$$
y_{ij} = \mu + \tau_i + \rho_j + \epsilon_{ij},
$$
$$
\epsilon_{ij} \sim N(0, \sigma^2),
$$
where $y_{ij}$ is the observation of the $i$th treatment in the $j$th block, $\mu$ is the overall mean, $\tau_i$ is the effect of the $i$th treatment, $\rho_j$ is the effect of the $j$th block, and $\epsilon_{ij}$ is the residual error.

# Exercise 2

2. Edit the R code

```{r}
library(tidyverse)
library(agridat)
```

```{r}
data("omer.sorghum")
```

```{r}
df <- omer.sorghum %>% filter(env == "E3")
model1 <- lm(yield ~ 1 + gen + rep, data = df)
```

```{r}
summary(model1)
```

```{r, echo=FALSE}
s2 <- sigma(model1)^2
```

The value for $\hat{\sigma}^2$ estimated in the model is `r format(round(s2,0), scientific = FALSE)`
